Stack Safety as a Security Property

### 2020-07-01 main.tex
Program and initial memory for example

Before we define stack safety, we need to first set the stage by
describing our machine model. Concretely, we have in mind a
conventional machine architecture with a standard (RISC-V) processor
extended with some kind of (static or dynamic, software- or
hardware-enforced) security monitor
\rb{Put stress on genericity first, advance this concrete model after?}.
However, our definition of stack
safety is generic: we will begin with an abstract interface to our
machine model and only make specific choices about details when we get
to the policy enforcement section~\ref{sec:micropolicy}.

@@

@@

@@

A trace over some type $A$ is a potentially infinite, nonempty
sequence of $A$s. We will mostly be concerned with sequences of
machines states, usually denoted $\machT$, and with sequences of
triples of (machine states, policy states, and observations), usually
denoted $\MPT$.
\rb{The way the definitions are set up, one may expect MPO, but\ldots}
To continue, we will require three \rb{four?} generic operations
on traces: a ``trace-of'' operation to construct the potentially
infinite traced induced by a step relation
\rb{not an op on traces per se?};
a ``split'', to break a
trace into two components based on a predicate; a ``join'', the
reverse of split; and a ``prefix'', as a shorthand to only keep the
first component of a split.

@@

@@

The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
and a predicate on elements \(f\) and gives the prefix of \(T\)
ending with the first element on which \(f\) holds, coinductively:

@@

@@

@@

@@

@@

@@

>>>>

Before we define stack safety, we need to first set the stage by
describing our machine model. Our definition of stack safety is
generic and this sections discusses an abstract interface to our
machine model and some kind of static or dynamic, software- or
hardware-enforced security monitor. At the same time, for
concreteness, we will build up our running example using a
conventional machine architecture with a standard (RISC-V) processor
and later extend it with a dynamic policy mechanism.

@@

\paragraph*{Running Example: RISC-V}

Mapping a conventional architecture to this abstract interface is
straightforward. To build up our running example, we are going to use
a standard 64-bit RISC machine, just like the one we are using in our
experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
r_{31}\} \cup {\PCname}$, and $\SP = r_2$. \leo{Mention We use the
  standard encoding of instructions to words?}

The first interesting choice is that of observations, as there are
multiple reasonable options. At one extreme, we could choose to define
observations as the value written at any component. However, that
would pose problems later on when we treat some components as ``secret''
when they correspond to a caller's private state. \leo{On the other extreme,
we could pick a particular component and treat it as the ``output'' of
our system. }

@@

\leo{TODO: Forward pointer to policies right?}

@@

A trace over some type $A$ is a potentially infinite, nonempty
sequence of $A$s. We will mostly be concerned with sequences of
machines states, usually denoted $\machT$, and with sequences of
triples of (machine states, policy states, and observations), usually
denoted $\MPT$.
\rb{The way the definitions are set up, one may expect MPO, but\ldots}
To continue, we will require four generic operations
on traces: a ``trace-of'' operation to construct the potentially
infinite traced induced by a step relation
\rb{not an op on traces per se?};
a ``split'', to break a
trace into two components based on a predicate; a ``join'', the
reverse of split; and a ``prefix'', as a shorthand to only keep the
first component of a split.

@@

\leo{TODO: Add an example of a trace once I figure out the notations}

@@

\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
  seems better}
The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
and a predicate on elements \(f\) and gives the prefix of \(T\)
ending with the first element on which \(f\) holds, coinductively:

@@

\subsection{Running Example}

\leo{TODO: Figure out presentation}

\newcommand*{\add}{\text{add}}
\newcommand*{\addi}{\text{addi}}
\newcommand*{\sw}{\text{sw}}
\newcommand*{\lw}{\text{lw}}
\newcommand*{\jal}{\text{jal}}
\newcommand*{\jalr}{\text{jalr}}
%\begin{lstlisting}
%START:
%  addi r2 r2 12 ; allocate 3 words in the stack
%  sw r2 42 -8; store 42 in a private location
%  sw r2 0 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 F ; call F and store return address
% 
%F:
%  addi r2 r2 4 ; allocate for ra 
%  sw r2 r1 -4 ; store the return address to sp-4
%  addi r2 r2 8 ; allocate for arg + ret
%  sw r2 17 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 G ; call G and store return address
% 
%G:  
%  addi r2 r2 4 ; allocate for ra 
%  sw r2 r1 -4 ; store the return address to sp-4
%  
%\end{lstlisting}

Starting State:

Instruction memory in locations 0 - ..
PC at 100.
Stack memory starting at 1000.
Assume we're in a function. 
Stack Pointer at 1040.
1040: 42.

\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-18.7em}{0.6em}{42}
\MemoryLabel{-16.4em}{0.6em}{0}
\\
\end{center}

Program:

\begin{verbatim}
top{
  return f(0);
}

f(x){
  return (x + g(17));
}

g(x){
  return (x + 1);
}
\end{verbatim}

\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
\newcommand*{\clabel}[1]{#1: & & &\\}
\newcommand*{\rsp}{sp}
\newcommand*{\rra}{ra}
\[
\begin{array}{llll}
  \clabel{TOP}  
  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
  & & & \\
  & & & \\
  \clabel{F}
  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
  & & & \\
  & & & \\  
  \clabel{G}
  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}

%  addi r2 r2 8 ; allocate for arg + ret
%  sw r2 17 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 G ; call G and store return address

\end{array}
\]

%  sw r2 42 -8; store 42 in a private location
%  sw r2 0 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 F ; call F and store return address

%  0: & \addi ~ r_4 ~ r_0 ~ 40 &
%     ; ~ \text{\em put $40$ in register $r_4$}\\
%  4: & \sw ~ r_4 ~ r_4 ~ 1000 &
%     ; ~ \text{\em store $40$ in location $1040$}\\
%  8: & \jal ~ r_1 ~ 20        &
%     ; ~ \text{\em jump to $20$, store return addres in $r_1$}\\
%  12: & & \\
%  16: & & \\
%  20: & \sw ~ r_2 ~ r_1 ~ 4 &
%     ; ~ \text{\em store the return address to the $\SP + 4$}\\
%  24: & \addi ~ r_2 ~ r_2 ~ 8 &
%     ; ~ \text{\em increase the $\SP$ by 8}\\
%\end{array}
%\]
% 
%  , (SW sp ra 4  , tagH1)
%  , (ADDI sp sp 8, tagH2)
%  , (LW 3 3 1000, instrTag)
%  ]

@@

\leo{Visible from assembly}

@@

\leo{The two JAL instructions map to 2 (arg + return)}

@@

\leo{retmap is the jal + 1}
\leo{entrymap is the ones after the labels}

@@

\leo{Point to the trace. two return points}

@@

  \leo{Snapshot state at the point of the call from TOP to F, collor
    the different things}
